---
title: "HW03"
author: "Nathan Bendriem"
date: "September 27, 2017"
output: 
  html_document: 
    keep_md: yes
---

Library(gapminder)
```{r}
library(gapminder)
library(tidyverse)
```

# Homework 3

For this homework, we will still be using the gapminder dataset, as well as dplyr and ggplot for graphs and tables


## We are given a list of tasks, and are asked to pick at least three

- Get the maximum and minimum GDP per capita within the continents
- look at the spread of GDP per capita within the conitnents
- Compute a trimmed mean of life expectancy for different years
- How is life expectancy changing over time on different continents
- Report the absolute and/or relative abundance of countries with low life expectancy
- Or make up your own!

## Task 1

I will start by answering the task "how is life expectancy changing over time on different continents"

To set up the dplyr table, we will be using the select function and the gapminder dataset.  I will only be focusing on the years 1957 and 2007, the first and last years available from the dataset, to illustrate the change in life expectancy


```{r}
A <- gapminder %>% group_by(country) %>% mutate(deltaLifeExp=lifeExp[year==2007]-lifeExp[year==1957]) %>% select(continent, country, deltaLifeExp)
knitr::kable(A)

```

The table shows us the country as well as the difference between life expectancy from 1957 to 2007.  

To graph this, I will use the facet_wrap function to indicate the 5 continents.  The graphs show the change in life expectancy over time (x-axis).  The geom_smooth function will add a regression line.

```{r}
ggplot(gapminder, aes(x=year, y=lifeExp)) + facet_wrap(~ continent) +  geom_point(aes(colour=continent)) + geom_smooth(se=FALSE) + labs(x= "Year", y="Life Expectancy") 
```


## Second Task

For the second task, I will be answering "Get the maximum and minimum of GDP per capita for all continents"

By using the group_by function, and defining the miniumum and maximum gdp per capita using the summarize function, we get a well organized table

G <- gapminder %>% group_by(continent) %>% summarize(min_gdpPercap=min(gdpPercap), max_gdpPercap=max(gdpPercap))

To graph this, I will be using a scatterplot to indicate the minimum and maximm gdp per capita.

```{r}
knitr::kable(G)
```
```{r}
ggplot(G, aes(x=continent, y=min_gdpPercap)) + geom_point(aes(colour=continent)) + labs(x="Continent", y="GDP Per Capita")
```



## Third Task

For the third task, I will compute a trimmed mean of the life expectancy.  First, I decided to filter out some countries.  We will be focusing only on Arctic countries.  We wil also focus on the time span of 1980-2000. 

I used the summarize function to define a new varialbe "mean_lifeExp" which is simply the average life expectancy of the countrys from the years 1980-2000.  I then used the arrange function with descending mean_lifeExp so that the table is organized with the countries that have the lowest average life Expectancy first

```{r}


Y <- gapminder %>% group_by(country) %>%  filter(year <= "2000", year >= "1980", country=="Norway" | country=="Sweden" | country=="Finland" | country=="Russia" | country=="United States" | country=="Canada" | country=="Denmark" | country=="Iceland") %>% summarize(mean_lifeExp=mean(lifeExp)) %>% arrange(mean_lifeExp, country)

knitr::kable(Y)
```
```{r}
ggplot(Y, aes(x=country, y=mean_lifeExp, colour=country)) + geom_point() + ggtitle("Average Life Expectancy of Arctic Countries (1980-2000)") + labs(x= "Country", y= "Life Expectancy")
```


## Fourth Task

For the final task, I will be looking at countries with low life expectancy and if GDP per capita is an indication of low life expectancy.

I define low life expectancy with an average of 45 years and below.

First I will identify those countries that have an average life expectancy of 45 years and below in the year 2007

```{r}
M <- gapminder %>% filter(year=="2007", lifeExp <= "45") %>% select(country, lifeExp, gdpPercap)

knitr::kable(M)
```

Then I will graph the Life Expectancy vs GDP per Capita and add a regression line to see which countries follow the trend and which do not.  

```{r}
ggplot(M, aes(x=lifeExp, y=gdpPercap)) + geom_point(aes(colour=country)) + labs(x="Life Expectancy", y="GDP Per Capita") + ggtitle("Life Expectancy vs GDP per Capita") 

```



