---
title: "HW05"
author: "Nathan Bendriem"
date: "October 14, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(gapminder)
library(forcats)
library(dplyr)
library(tidyverse)
library(knitr)
library(tidyr)
```


# WELCOME! to the fifth and final homework assignment for STAT545

The goals for this homework are:
- Reorder a fctor in a principled way based on the data and demonstrate the effect in arranged data and figures
- Improve a figure using new knowledge
- Implement visualtion design principles
-  Write some data to file and load it back into R
- Organize my Github as a final goodbye 

Lets get right into it!

## Factor Management

To go full circle, I decided to go back to the gapminder dataset since we've been focusing a lot on the singer dataframes in class lately.

Step goals:
- Define factor variables
- Drop factor/levels
- Reorder levels based on knowledge from data

### Drop Oceania

```{r}
Oceania1 <- gapminder %>% filter(continent=="Oceania")
```

Oceania is a continent with two countries (Australia and New Zealand) with 24 observations in the gapminder dataset

```{r}

```


```{r}
glimpse(Oceania)
```

```{r}
Bye_Oceania <- gapminder %>% mutate(oceania_factor=as_factor(oceania) %>% 
```

