---
title: "HW05"
author: "Nathan Bendriem"
date: "October 14, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(RColorBrewer))
```




# WELCOME! to the fifth and final homework assignment for STAT545

The goals for this homework are:
- Reorder a fctor in a principled way based on the data and demonstrate the effect in arranged data and figures
- Improve a figure using new knowledge
- Implement visualtion design principles
-  Write some data to file and load it back into R
- Organize my Github as a final goodbye 

Lets get right into it!

## Factor Management

To go full circle, I decided to go back to the gapminder dataset since we've been focusing a lot on the singer dataframes in class lately.

Step goals:
- Define factor variables
- Drop factor/levels
- Reorder levels based on knowledge from data

### Drop Oceania

First lets look at the original gapminder dataframe

```{r}
nrow(gapminder)
levels(gapminder$continent)
```

and lets look at the number of observations in Oceania

```{r}
Oceania <- gapminder %>% filter(continent=="Oceania")

str(Oceania)
```

Oceania has 24 observations in the gapminder dataset

Next we will drop the factor Oceania

```{r}
ByeOceania <- gapminder %>% 
  filter(continent != "Oceania") %>% 
  droplevels()
```

Now lets check the numbers of rows in Gapminder, it should be 24 observations less than the original gapminder

```{r}
nrow(ByeOceania)
levels(ByeOceania$continent)
str(ByeOceania)
forcats::fct_count(ByeOceania$continent)
```


### Factor Reorder

Reorder the levels of country.  Use the forcats package to change the order of the factor levels baes on a principled summary of one of the quantiatve variables.

The variable I will be working with is the gdp Per capita.  We will be focusing on the countries with the maximum gdp per capita.  By using the head() function, we only keep the top 6 countries with the highest gdp per capita.

```{r}
ContinentReOrder <- fct_reorder(ByeOceania$country, ByeOceania$gdpPercap, max, .desc= TRUE) %>% 
  levels() %>% 
  head()
```

With the reordering done, I will define MaxGDP as a dataframe with only the year, country, and the gdp per capita.  It eliminates the life Expectancy and the continent factors

```{r}
MaxGDP <- ByeOceania %>% 
  filter(country %in% ContinentReOrder) %>% 
  select(year, country, gdpPercap)
```

Lets plot Max GDP!

```{r}
ggplot(MaxGDP, aes(year, gdpPercap, colour= country)) + facet_wrap(~country) + 
  geom_line() + scale_colour_brewer(palette="Dark2") + 
  labs(x= "Year", y= "GDP Per Cap", title="Countries with Max GDP per Cap") + theme_bw()
```

### Common Part

Characterieze the derived data before and after your factor re-leveling


In the table below, you'll notice that it is arrangd in alphabetical order by country, not by descending Maximum GDP per capita

```{r}
MaxGDP %>% group_by(country) %>% 
  knitr::kable(format="html")
```



And the plot looks like this:


```{r}
ggplot(MaxGDP, aes(year, gdpPercap, colour= country)) + facet_wrap(~country) + geom_line() + 
  scale_colour_brewer(palette="Dark2") + 
  labs(x= "Year", y= "GDP Per Cap", title="Countries with Max GDP per Cap") + theme_bw()
```

Now we re-arrange the dataframe to descend in terms of GDP per Cap

```{r}
AsceMaxGDP <- MaxGDP %>% 
  group_by(country) %>% 
  arrange(desc(gdpPercap))
knitr::kable(AsceMaxGDP, format="html")
```

and the plot looks like this:

```{r}
ggplot(AsceMaxGDP, aes(year, gdpPercap, colour= country)) + facet_wrap(~country) + 
  geom_line() + 
  scale_colour_brewer(palette="Dark2") + 
  labs(x= "Year", y= "GDP Per Cap", title="Countries with Max GDP per Cap") + theme_bw()
```

Yet on the graph, the order is still aphabetical rather than by the country with the highest gdp Per capita.  

We are going to attempt to change the graph so that it is ordered by descending gdp per cap

```{r}
HighGDP <- MaxGDP %>% 
  group_by(country, year) %>% 
  summarize(maximumgdp= max(gdpPercap)) %>% 
  arrange(desc(maximumgdp))

knitr::kable(HighGDP, format="html")
```


In this graph, the legend illustrates the country with the highest gdp in descending order:


```{r}
ggplot(HighGDP, aes(year, maximumgdp, colour= fct_reorder(country, maximumgdp, .desc=TRUE))) + 
  facet_wrap(~country) + geom_line() + scale_colour_discrete(name="Highest GDP per Cap") + 
  labs(x= "Year", y= "GDP Per Cap", title="Countries with Max GDP per Cap") + theme_bw()
```

### File I/O

Experiment with write_csv or read_csv.  

```{r}
write_csv(MaxGDP, "~/Lets-Practice-CSV")
MaxGDP_Read <- read_csv("~/Lets-Practice-CSV")
```


## Visualization Design

For this section of the homework, I will take an old figure I made in Homework 2 and spice it up.  I'm going to change the color scheme, the code chunk, and add a legend.

First, here is the graph and code as it was written in [HW2](https://github.com/nbendriem/STAT545-HW-bendriem-nathan/blob/master/STAT545HW2/Answers_to_HW02.md)

```{r}
NorthAfrica <- filter(gapminder, country=="Algeria" | country=="Tunisia" | country=="Egypt" | country=="Morocco" | country=="Libya" | country=="Sudan")

ggplot(NorthAfrica, aes(x= country, y=lifeExp)) + 
  geom_boxplot(colour= "black", fill="red", alpha=0.33) + labs(x="Country", y="Life Expectancy") + 
  ggtitle("Boxplot of Life Expectancy of North African Countries")

```

I am going to start by cleaning up the code chunk

```{r}
NorthAfrica2 <- filter( gapminder, country==c("Algeria", "Tunisia", "Egypt", "Morocco", "Libya", "Sudan"))

```

And the graph as well


```{r}
NAfricaPlot <- ggplot(NorthAfrica2, aes(country, lifeExp)) + geom_boxplot(aes(fill=country)) + 
  scale_fill_brewer(palette = "Blues") + 
  labs(x="Country", y="Life Expectancy", title="Life Expectancy of N. African Countries") + guides(fill=guide_legend(title="North African \nCountry")) + 
  theme_bw() 

NAfricaPlot
```


## Writing figures to file

I will use ggsave to save the plot above and then load and embed it in this report

```{r}
ggsave("My_Plot.pdf", NAfricaPlot, width=6, height=4, units="in")
```

And this plot is saved within the Homework 5 folder!

to embed it to this markdown file:


![](My_Plot.pdf)


