---
title: "HW04"
author: "Nathan Bendriem"
date: "October 6, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)
library(gapminder)
library(tidyr)
library(knitr)


## Homework 4!


This week's homework is another "Pick Your Own Adventure" type where we will be working on our data wrangling skills with some aggregation and reshaping functions.

One of the prompts focuses on data reshaping while the second prompt while  focus on the join functions we have learned last week

### General Data Reshaping and relationship to aggregation

I will be answering the second activity: make a tibble with one row per year and columns for life expectancy for two or more countries.

Lets take the 5 countries that are closest to Antarctica and compare their life expectancies.


```{r}
antarctic_countries <- spread(select(filter(gapminder, country=="Chile" | country=="Argentina" | country=="South Africa" | country=="Australia" | country=="New Zealand"), year, country, lifeExp), country, lifeExp)
```



Here I used the spread function to create a table of 6 Variables measuring life Expectancy for each year listed in the Gapminder Dataset

I will use tidyr and kable functions to clean up the table, and add a title


```{r}
knitr::kable(antarctic_countries2, format="markdown", col.names=c("Year", "Argentina", "Australia", "Chile", "New Zealand", "South Africa"), caption="Life Expectancy of Antarctic Countries")
```
Next, we will plot the life Expectancy of Argentina vs that of Chile.  I used the scale_colour_discrete function to take out the title of the legend and rename the dot from "red" to "Life Expectancy"

```{r}
ggplot(antarctic_countries2) + geom_point(aes(x=Chile, y=Argentina, colour= "Red")) + ggtitle("Life Expectancy of Argentina vs Chile") + scale_colour_discrete(name=NULL, labels=c("Life Expectancy"))
```

### Join, Merge, Look Up

I created two new dataframes with a couple of shark species found in the Pacific Ocean.  
```{r}
setwd("~/Desktop/Stat545: R/STAT545-HWs-Bendriem-Nathan/STAT545HW4")

S_LE <- read.csv("~/Desktop/Stat545: R/STAT545-HWs-Bendriem-Nathan/STAT545HW4/shark_lifeExp.csv")

S_Loc <- read.csv("~/Desktop/Stat545: R/STAT545-HWs-Bendriem-Nathan/STAT545HW4/shark_location.csv")

```

The first dataframe has species, length in meters and life expectancy

```{r}
kable(S_LE)
```

The second dataframe has a couple of the same species, and where they are found, whether its coastal, pelagic, or benthic

```{r}
kable(S_Loc)
```

Then I play around with the join functions with both data frames. 

# Left and Right Join

Left join will take matching rows from the second variable and add them to the first variable.  Right join will do the inverse and add matching rows from the first variable and add them to the second variable

```{r}
left_join(S_LE, S_Loc, by="Species") %>% kable()
```
Here, all species of sharks are presented and those the corresponding location for the sharks found in S_LE and not in S_Loc come up as NA
```{r}
right_join(S_LE, S_Loc, by="Species") %>% kable()
```

As you can see, in the right join, only the species that are found in the S_Loc dataframe are depicted.  


# Inner and Full Join

Inner and Full join will work with rows instead of columns.  Rows with that match from both dataframes will be retained with inner join. Full join will retain all rows.  


```{r}
inner_join(S_LE, S_Loc, by="Species") %>% kable()
```
For inner_join, only the species of sharks found in both dataframes are shown

```{r}
full_join(S_LE, S_Loc, by="Species") %>% kable()
```
Here all species of sharks are shown

# Semi and Anti Join

Semi join does not actually join one column to another data frame but will filter out all rows in the first dataframe that have a match in the second data frame.  Anti join does the opposite and only shows the rows that in the first dataframe that are not matched in the second dataframe

```{r}
semi_join(S_LE, S_Loc, by="Species") %>% kable()
```
Here only the sharks found in both dataframes are shown, but since S_LE was the first dataframe, its columns are retained

If we inverse the two dataframes we get this:

```{r}
semi_join(S_Loc, S_LE, by="Species") %>% kable()
```
Once again, only the sharks found in both dataframes are shown, but the column comes from the S_Loc dataframe since it is the first variable


For the anti_join function, only the sharks that are not found in both databases are depicted
```{r}
anti_join(S_LE, S_Loc, by="Species") %>% kable()
```

# Binding

The binding function will addd either rows (bind_rows) or columns (bind_columns) into a new dataframe

```{r}
bind_rows(S_LE, S_Loc) %>% kable()
```
 
```{r}
bind_cols(S_LE, S_Loc) %>% kable()
```
bind columns does not work because the number of rows from each dataframe are not the same.  One has 14, one has 8








