---
title: "Hw4 pt 4"
author: "Nathan Bendriem"
date: "October 9, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)
library(gapminder)
library(tidyr)
library(knitr)

### Join, Merge, Look Up

I created a new data frame of capital cities to complement the countries using Excel and saving it as a .csv file.  

read.csv("~STAT545-HW-bendriem-nathan/STAT545HW4/capital cities2.csv")
```{r}
setwd("~/Desktop/Stat545: R/STAT545-HWs-Bendriem-Nathan/STAT545HW4")




```

```{r}
CC <- read.csv("~/Desktop/Stat545: R/STAT545-HWs-Bendriem-Nathan/STAT545HW4/capital cities3.csv")
```

Then I play around with the join functions with the Gapminder and capital cities data frames. 

# Left and Right Join

Left join will take matching rows from the second variable and add them to the first variable.  Right join will do the inverse and add matching rows from the first variable and add them to the second variable

```{r}
left_join(gapminder, CC) %>% kable()
```


```{r}
right_join(gapminder, CC) %>% filter(year=="2007") %>% kable()

```


# Inner and Full Join

Inner and Full join will work with rows instead of columns.  Rows with that match from both dataframes will be retained with inner join. Full join will retain all rows.  


```{r}
inner_join(gapminder, CC) %>% filter(year=="1957") %>% kable()
```
```{r}
full_join(gapminder, CC) %>%  filter(year=="2007") %>% kable()
```


# Semi and Anti Join

Semi join does not actually join one column to another data frame but will filter out all rows in the first dataframe that have a match in the second data frame.  Anti join does the opposite and only shows the rows that in the first dataframe that are not matched in the second dataframe

```{r}
semi_join(gapminder, CC) %>% filter(year=="1957") %>% kable()
```
```{r}
anti_join(gapminder, CC) %>% filter(year=="1957") %>% kable()
```
